---
title: "Reproducibility Report: Maglio & Polman (2014), Study 1"
subtitle: "Minimal version"
author: "Your name here"
date: today
format:
  html:
    toc: true
    toc-location: left
    embed-resources: true
execute:
  error: true
---

:::{.callout-note title="Goals"}
1. Load a published dataset with `here()` and `read_excel()`
2. Reproduce the ANOVA and effect sizes reported in the paper
3. Say clearly what did and did not reproduce, and why
:::

This is the **minimal** version: the structure and the questions are here, and
you write the code. If you get stuck, `reproducibility-report-bumper-rails.qmd`
has the same steps with most of the code filled in. It is fine to peek at it
for one step and then come back.

`error: true` is set at the top of this file, so the document still renders
while some chunks are unfinished. The chunks that still fail show their error
message in red.

Before you start, read Study 1 in the paper:
`original_paper/maglio-polman-2014.pdf` (two pages).

## The study

Maglio, S. J., & Polman, E. (2014). Spatial orientation shrinks and expands
psychological distance. *Psychological Science, 25*(7), 1345–1352.
<https://doi.org/10.1177/0956797614530571>

**Methods summary.** Researchers recruited 202 volunteers at the Bay Street
subway station in Toronto. Half were waiting on the westbound platform and half
on the eastbound platform. Each participant was randomly assigned to rate how far
away one of four stations felt (1 = very close, 7 = very far): Spadina (two stops
west), St. George (one stop west), Bloor-Yonge (one stop east), or Sherbourne
(two stops east). The authors ran a 2 × 4 ANOVA on the distance ratings, then
compared eastbound and westbound travelers separately at each station.

**Target outcomes.** This is the result you are trying to reproduce, quoted from
the Study 1 results section:

> We carried out a 2 (orientation: toward, away from) × 4 (station: Spadina, St.
> George, Bloor-Yonge, Sherbourne) analysis of variance (ANOVA) on closeness
> ratings, which revealed no main effect of orientation, F < 1, and a main effect
> of station, F(3, 194) = 24.10, p < .001, ηp² = .27. This main effect was
> qualified by the predicted interaction between orientation and station, F(3,
> 194) = 16.28, p < .001, ηp² = .20. We decomposed this interaction by the
> subjective-distance ratings between participants traveling east and west for
> each of the four subway stations. Westbound participants rated the stations to
> the west of Bay Street as closer than did eastbound participants; this effect
> was obtained for both the station one stop to the west (St. George, p < .001,
> ηp² = .28) and the station two stops to the west (Spadina, p = .001, ηp² =
> .20). The opposite pattern held true for stations to the east of Bay Street.
> Eastbound participants rated the stations to the east of Bay Street as closer
> than did westbound participants; this effect was obtained for both the station
> one stop to the east (Bloor-Yonge, p = .053, ηp² = .08) and the station two
> stops to the east (Sherbourne, p < .001, ηp² = .24). Figure 1 summarizes these
> results.

ηp² is **partial eta squared**, an effect size. You will compute it below.

## Step 1: Load packages

`readxl` (reads Excel files) and `broom` (turns model output into tibbles) are
installed along with the tidyverse, but `library(tidyverse)` does not load them,
so we load them by name.

```{r}
#| eval: false
install.packages(c("tidyverse", "here"))
```

```{r}
#| message: false
library(tidyverse)
library(readxl)
library(broom)
library(here)
```

## Step 2: Load the data

The authors posted one Excel file per study. We only need Study 1. The columns
are described in `docs/codebook.md`.

```{r}
d <- read_excel(here("data", "S1_Subway.xlsx"))

glimpse(d)
```

## Step 3: Tidy the data

The data are already tidy: one row per participant, one column per variable.
But R has read `DIRECTION` and `STN_NAME` in as text. For an ANOVA they need to
be **factors**, and we set the order of the station levels to run west to east,
so tables and plots come out in map order.

Use `mutate()` and `factor()` to convert both columns. For `STN_NAME`, give
`factor()` a `levels` argument listing the stations from west to east.

```{r}

```

## Step 4: Reproduce the analysis

### How many people are in each cell?

Count the participants in each direction × station combination. Does it match
the paper (202 total, 101 per platform)? Is every cell the same size?

```{r}

```

### The 2 × 4 ANOVA

Fit a two-way ANOVA with `aov()`: distance predicted by direction, station, and
their interaction. Remember that `y ~ A * B` is shorthand for `y ~ A + B + A:B`.

```{r}

```

The numbers in parentheses in "F(3, 194)" are the degrees of freedom: 3 for the
effect, 194 for the residuals. Check that yours match.

The teaching team gets:

- Main effect of direction: F(1, 194) = 0.66
- Main effect of station: F(3, 194) = 23.35
- Interaction between direction and station: F(3, 194) = 16.28

Two of those match the paper. One is close but not the same. Keep going; the
bonus question at the end asks you to work out why.

### Partial eta squared

Partial eta squared is the share of variance an effect accounts for, once
the other effects are set aside:

$$
\eta_p^2 = \frac{SS_\text{effect}}{SS_\text{effect} + SS_\text{residual}}
$$

Compute it for each effect. Hint: `tidy()` from broom turns the ANOVA table into
a tibble, with the sums of squares in a column called `sumsq`.

```{r}

```

Do your values round to the paper's .27 (station) and .20 (interaction)?

### Station-by-station comparisons

Now compare eastbound and westbound travelers at each station separately. For
each station: keep only that station's rows, compute the mean rating in each
direction, run a one-way ANOVA of distance on direction, and compute ηp².

#### St. George

```{r}
#| label: st-george

```

#### Spadina

```{r}
#| label: spadina

```

#### Bloor-Yonge

```{r}
#| label: bloor-yonge

```

#### Sherbourne

```{r}
#| label: sherbourne

```

Fill in this table as you go:

| Station | Paper *p* | Your *p* | Paper ηp² | Your ηp² |
|---|---|---|---|---|
| St. George | < .001 | | .28 | |
| Spadina | .001 | | .20 | |
| Bloor-Yonge | .053 | | .08 | |
| Sherbourne | < .001 | | .24 | |

## Step 5: Reflection

Were you able to reproduce the results you attempted to reproduce? If not, which
parts could you not reproduce?

> ANSWER HERE

How difficult was it to reproduce the results?

> ANSWER HERE

What made it hard, and what made it easy? Think about the data files, the
variable names, and how much the paper told you about the analysis.

> ANSWER HERE

## Bonus questions

**1. Why is the station F different?** Your F for station (23.35) is close to
the paper's (24.10) but not equal, while the interaction matches exactly. Try
refitting the model with the predictors in the opposite order,
`DISTANCE ~ STN_NAME * DIRECTION`. Does anything change? Look back at your cell
counts. What does that suggest about how the authors' software (SPSS) and R's
`aov()` each divide up the variance? `docs/anova-in-r.md` explains it, but try
first.

```{r}

```

**2. Orientation or direction?** The paper describes the factor as
"orientation: toward, away from". The dataset has `DIRECTION`: east or west.
Are those the same variable? If not, which one did the authors actually put in
their ANOVA? Make an `orientation` column with `if_else()` and refit the model
with it to check.

```{r}

```

**3. Reproduce Figure 1.** Plot the mean distance rating for each station,
split by direction, with error bars of ±1 standard error, as in the paper.
`stat_summary(fun.data = mean_se)` will draw both the means and the error bars
for you.

```{r}

```

## Going further

Instead of four near-identical chunks, do all four stations in one pipeline.
Look up `nest()` and `map()`.

```{r}

```
